---
title: My First Post!
author: Zachary Hamilton
date: '2018-10-26'
slug: my-first-blog-post
categories: []
tags: ["R", "Baseball", "Data Vis"]
---

## Intro 

Hi, my name is Zach. Welcome to my blog, which I have created to document my journey into the world of _Data Science_. Playing with data and statistics/probability has always been the thing I could do for hours and hours and not realize that any time had passed. I began learning R so that I could dive deeper into this passion of mine. I chose R for the array of open source learning materials and how accessible the community felt. R is also incredibly powerful and friendly to novice programmers (think learning how to ride a bike and doing backflips off of jumps at the end of the first day). I plan to use this outlet to practice my data wrangling and analysis skills in R on datasets that span my many interests.  

With that said, here is the first plot that I ever made in R. I tracked the weight of my labrador puppy, Olive, over the first few months of her life and used R to visualize this trend:

```{r Olive, message = FALSE, warning = FALSE, echo = FALSE}
#load packages
library(tidyverse)
library(lubridate)

setwd("C:/Users/Zachary/Documents/R/data")

# read in data 
dog_weight <- as.tibble(read.table("dog_weight.txt", 
                                   col.names = c("date", "weight"),
                                   sep = ":", skip = 1))
# get R to treat date as a date
dog_weight <- dog_weight %>%
  mutate(date = mdy(date))

#plot time 
ggplot(dog_weight, aes(date, weight)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  labs(title = "Olive's Weight Chart")

setwd("C:/Users/Zachary/Documents/R/Projects/blog")
```

I had just finished reading [R for Data Science](https://r4ds.had.co.nz/) and had a very vague knowledge of tidyverse data manipulation verbs and knew that they played well with the __ggplot__ function so this was the result (I hadn't learned much about formatting yet...) It might not be much, but I realized from this simple exercise how powerful R could be and I absolutely had to learn more.

After that, I did some more reading and became very excited by the concept of the _pipe operator_, as you can see from the code below, and came up with a visualization that shows the batting average of AL West Division MLB teams from 2011 to 2016, courtesy of data from the Lahman database!

```{r baseball, message=FALSE, warning=FALSE}
library(tidyverse)
library(Lahman)

#Batting average by team 

Batting %>%
  filter(yearID > 2010, lgID == "AL") %>%
  left_join(Teams, by = c("yearID", "teamID")) %>%
  filter(divID == "W" & lgID.x == "AL") %>%
  group_by(yearID, teamID) %>%
  summarise(batavg = round(sum(H.x) / sum(AB.x), 3)) %>%
  ggplot(aes(yearID, batavg, colour = teamID)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0.3) +
  ggtitle("Batting Average") +
  labs(x = "Year", y = "Batting Average")
```

I won't normally show my code unless I am illustrating a certain new concept I've learned or something, but all source material will be available on my [Github](https://github.com/zjham/blog). Big thanks to Jenny Bryan's [helpful resource](http://happygitwithr.com/) on Git & version control and Yihui Xie's [intro to blogdown](https://bookdown.org/yihui/blogdown/) for helping me get this blog up and running painlessly. I'll end here for now, I promise the content in future posts will have nicer formatting. Thanks for reading! 


